### 2015-04-19 summary.tex
minor edits to intro

\section{Summary and discussion}\label{summary}
In this paper we have studied the axisymmetric vertical shear 
instability (VSI) in astrophysical disks that arise from the 
height dependence of the equilibrium angular velocity profile 
$\Omega=\Omega(r,z)$ in a baroclinic disk. Vertical shear is generally
expected in astrophysical disks due to the radial variation in the
disk temperature, or more generally the entropy. However, 
astrophysical disks are also typically sub-adiabatically stratified. 
In this case, the VSI requires a short thermal timescale $t_c$ in
order to operate.

Physically, a short thermal timescale is needed to reduce the  
stabilizing effect of buoyancy forces in a stably stratified
disk. Because vertical shear is weak in thin disks, $t_c\Omega\ll 1$
is generally needed for the VSI to operate, otherwise buoyancy would
be strongly stabilizing. Our primary goal was to quantify this
requirement, thereby provide a simple way to assess 
whether or not the VSI is relevant in realistic disks.

To do so, we generalized previous vertically global linear 
calculations of the VSI \citepalias{nelson13,mcnally14,barker15}, which
assumed isothermal perturbations, to include an energy equation with 
a source term in that relaxes the disk
temperature to its initial value on a timescale
$t_c$. This `thermal relaxation' model allows
the thermodynamic response of the disk to be controlled by varying
$t_c$, and has been employed in previous numerical simulations of the
VSI \citepalias{nelson13}.

In accordance with the above studies, which show that the VSI occurs for 
perturbations with small radial lengthscales, we applied the 
radially local approximation in the linearized fluid equations. 
We mostly focused on vertically isothermal disks as 
this is the relevant case for protoplanetary disks (PPDs) passively
irradiated by its central star \citep{chiang97}. It also permits a
largely analytical discussion. Our analysis is then an extension of the
linear models considered in \cite{lubow93} by adding vertical shear
and (rapid) thermal relaxation. We obtained explicit solutions 
describing the VSI with thermal relaxation in the thin disk limit.

For a vertically isothermal, sub-adiabatically stratified disk we
determined the thermal timescale   
\begin{align*}
  t_\mathrm{crit} = \frac{\epsilon |q|}{\gamma-1}\Omega_k^{-1}, 
\end{align*}
below which the VSI operates effectively. While $t_\mathrm{crit}$ was
derived for the lowest order VSI mode --- the `fundamental  
corrugation mode' --- we find from numerical calculations that
$t_\mathrm{crit}$ can be used to characterize the general
stabilization of the VSI by a finite thermal timescale. %That is, all
%VSI growth rates are reduced significantly as $\beta$ is increased
%from 
%zero to $\beta_\mathrm{crit}$.    
We have therefore provided a quantitative measure for the thermal
timescale requirement for the VSI in astrophysical 
disks. The thermal timescale must be shorter than the disk
dynamical timescale by $O(\epsilon)$, which is
quite short for thin disks ($\epsilon\ll 1$), because vertical shear is weak.

We applied our linear models to assess the applicability of the VSI in
PPDs. We considered the Minimum Mass Solar Nebula 
(MMSN) described in \cite{chiang10} and   
modeled the thermal timescale $t_c$ based on realistic
disk properties and dust opacity. 
%This results in $t_c$ being a
% function of position and the perturbation lengthscale of interest. 
We find the MMSN is unstable to the VSI 
from $\sim 5$AU to $\sim50$AU with characteristic growth times of
$\sim 30$ orbits. It is possible to have the VSI at smaller
radii, but only on radial lengthscales signficantly shorter
than the disk scaleheight, and such perturbations may
be subject to viscous decay in a real disk.

Although $t_c$ is not vertically constant in the 
MMSN as assumed in our analysis, we find that evaluating the condition
$t_c< t_\mathrm{crit}$ at the disk midplane gives an indication of the
radii at which the VSI operates in the MMSN. This simple criterion is
thus a practical first step in estimating the relevance of the VSI in
realistic astrophysical disks without resorting to linear
calculations or non-linear numerical simulations.

\subsection{Caveats and outlooks} 
We highlight below some extensions to the present  
linear models in future works.

\subsubsection{Convective overstability}
We have focused on parameter regimes relevant to the VSI,
namely perturbations with small radial lengthscales 
($k_xH\gtrsim 1$) in disks with short thermal timescales
($t_c\Omega\ll 1$). This excludes the `convective overstability'
recently discussed by \cite{klahr14} and \cite{lyra14} in vertically
unstratified disk models.

The convective overstability corresponds to slowly growing epicyclic motions.
It is most effective for $t_c\Omega \sim 1$ and disturbances with
$|k_x|\ll |k_z|$, where $k_z$ is a vertical wavenumber. These
conditions are mutually exclusive with the VSI.  
The maximum growth rate for convective overstability is given by 
\begin{align}
  \nu_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
\end{align}
\citep{lyra14}, where $N_r^2 = -(c_p\rho)^{-1}\p_rP\p_rS$ is the
square of the radial buoyancy frequency
(cf. Eq. \ref{nzsq_def}). Thus, growth requires $N_r^2<0$.  
%Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
%we find  
%\begin{align}
%  \nu_\mathrm{conv,max} =
%  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
%  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
%\end{align}
In fact, for the MMSN considered in this work, convective
overstability does not apply because 
$N_r^2>0$.

For more general disk structures convective overstability is
possible in principle. The models adopted for our numerical study in
\S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
require perturbations with long radial lengthscales,
$k_xH\ll 1$, for which radially global disk models would be more 
self-consistent.

\subsubsection{Improved thermal evolution} 
We assumed non-adiabatic effects can be represented by a thermal
relaxation timescale $t_c$. For  
application to PPDs we connected $t_c$ to the disk
properties such as surface density, temperature and opacity, making
$t_c$ depend on position and the perturbation 
lengthscales. One might improve the realism further by generalizing
the opacity model adopted here to be density (and hence height)
dependent.  
%here to be density-dependent,  i.e. $\kappa_d =
%% \kappa_d(\rho, T)$. 
However, going beyond thermal relaxation requires a
proper treatment of radiative diffusion. % The linear problem then
% involves a Laplacian operator and additional boundary
% conditions must be supplemented. 
We will consider these improvements in a
follow-up study. %additional heat source

\subsubsection{Additional physics} 
Depending on where the VSI operates in a
disk, additional effects may become important. For example, in the
outer parts of a PPD, where we estimate the VSI 
operates, disk self-gravity (SG) may play a role. Vertical
self-gravity may favor the VSI by increasing $|\p_z\Omega|$ through
enhancing the density stratification 
(Eq. \ref{vertical_shear}). %sg in pert? no because kx>>1. but maybe for kx~1

We have neglected viscosity. \citetalias{nelson13} find that an alpha
viscosity $\lesssim O(10^{-4})$ is required for the
VSI. \citetalias{barker15} found that inclusion of viscosity is needed for 
convergence when studying surface modes. %but not body modes? 
Viscosity may also affect the
thermal timescale requirement of the VSI, because viscous heating
needs to be self-consistently included in the energy equation. This
should be investigated in future works.

% It is of interest to quantify the effect of viscosity 
% on the linear VSI. This would help identify, more accurately, the
% appropriate regions in a protoplanetary disk that are sufficiently
% laminar to develop the VSI.  This problem is, however, complicated by
% the fact that proper inclusion of viscosity introduces a meridional
% flow in the equilibrium state, as well as viscous heating
% in the energy equation.