### 2015-04-23 linear_iso.tex
messed up notation change. need to fix by inspection.

\subsection{Shearing box, low-frequency and Keplerian
  approximations}\label{approx_gov}

In Appendix \ref{adia_improve} we derive a governing equation with the following simplifications:

\begin{enumerate}

\item  We adopt the  `vertically global shearing box' (VGSB) formalism 
  developed by \citetalias{mcnally14}, in which explicit dependencies on the
  background radial disk structure in 
  Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
  $\hat{g}_c=0$. However, the global radial temperature 
  gradient in the expression for the 
  vertical shear (Eq. \ref{vertical_shear}) is retained.

  In Appendix \ref{global_corr}  we discuss limitations of using
  the VGSB, where we show that this approximation is valid provided we 
  consider consider fast thermal relaxation times
  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
  ($k_xH\gtrsim 1$), as appropriate for the VSI.

\item We consider eigenvalues such that
  $|\zeta^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency
  approximation. We are therefore considering inertial waves
  \citepalias{barker15}.

\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
  0$. This is already captured explicitly in the linearized equations through
  the term $d\Omega/dz$. We therefore simplify 
  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
  elsewhere. For clarity, we replace $\kappa^2$ with
  $\Omega_K^2$. We call this the Keplerian approximation.

\end{enumerate}
% Making both of these approximations is equivalent to setting 
% $D\to\Omega_K^2$.  After making this replacement, we eliminate $W$ and
% $Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
% making use of Eq. \ref{vertical_shear}, to
% obtain an equation for $\delta v_z$.

These approximations lead to Eq. \ref{full_ode} which, for a thin disk
$h\ll 1$, simplifies further to  
%The result for $\Gamma=1$ is
\begin{align}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
      q}{\Omega_K^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  % &+ \left[\zeta^2\left(\frac{k_x^2}{\Omega_K^2} +
  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
  &+ \left[\frac{\zeta^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{\Omega_K^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
  &\phantom{+=}\left.-
    \frac{c_s^2}{\Omega_K^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
    \left(1-\chi\right) 
  \right]\delta v_z,\label{vertiso_gov}
\end{align}
where
\begin{align}
  \chi \equiv \frac{1-\ii\zeta t_c}{1-\ii\zeta t_c\gamma}.
\end{align}

It is convenient to introduce the dimensionless variables
\begin{align}
  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\zeta} =\hat{\omega} +
  \ii\hat{\sigma}= \zeta/\Omega_K,
\end{align}
where $\hat{\omega}$ and $\hat{\sigma}$ are real. 
Eq. \ref{vertiso_gov} becomes 
\begin{align}
  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii h
    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
  &+
  \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
    \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}\right.
  \notag\\
  &\left.- \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii h
      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
\end{align}
where $^\prime$ denotes $d/d\zhat$.

\subsection{Thin-disk approximation}\label{analytic_relax}
Our final approximation apply to thin disks for 
which $h\ll 1$. Then we may write $\rho/\rho_0\simeq
\exp{(-\zhat^2/2)}$. The governing equation,
Eq. \ref{vertiso_gov_nondim}, is then   
\begin{align}
  \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
  (B - C\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
\end{align}
with
\begin{align}
  &A \equiv 1 + \ii h q \hat{k},\\
  &B \equiv \hat{\zeta}^2\left(\chi + \hat{k}^2\right) -
  \left(\chi + \ii h q \hat{k}\right),\\
  &C \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
    h q\hat{k}\right), 
\end{align}
and $\chi =
\left(1-\ii\hat{\zeta}\beta\right)/\left(1-\ii\hat{\zeta}\beta\gamma\right)
$ in terms of dimensionless variables.

We remark that Eq. \ref{nearly_iso_explicit} is in the same 
form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
a vertically isothermal disk without vertical shear. The two equations
become identical after  setting $q=0$ (no vertical shear) and
$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
and making the low-frequency, Keplerian approximation in
\citeauthor{lubow93} after a change of variables.

\section{Analytical discussion of vertically isothermal
  disks in the framework of the vertically global shearing
  box}\label{analytical}  
We begin with a discussion of vertically isothermal disks
where $\Gamma=1$, then $c_s$ is a constant in the radially-local
approximation. 
In order to facilitate a fully analytic discussion, we first introduce a
number of approximations below that allows a simple governing equation
to be derived and compared with previous works. Similar approximations
are employed by \citetalias{nelson13} and \citetalias{barker15}. However, all of
the approximations described below are relaxed in our numerical study.

%vertically isothermal disk lacks a surface

% \subsection{Relation to the vertically global
%   shearing box}\label{vgsb_approx} 
% We connect the linearized equations 
% Eq. \ref{lin_mass}---\ref{lin_energy} to the vertically global  
% shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
% is an extension of the standard shearing box \citep{goldreich65} to 
% background shear flows that are height dependent. 
% As in the  standard shearing box, in the VGSB the equilibrium disk
% structure is assumed to be homogeneous in the radial ($x$) direction.

% Thus, the linear problem in the VGSB can be derived from
% Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
% $\hat{g}_c=0$. The resulting problem is very similar 
% to that for local axisymmetric waves in accretion disks already
% discussed by \cite{lubow93}, but modified by the presence of vertical
% shear and thermal relaxation. This is convenient as it allows us to
% directly apply some of their solution methods.

% After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
% \ref{lin_energy}, we can eliminate variables 
% to obtain a pair or ordinary differential
% equations for $W,Q$ as
% \begin{align}
%   &\frac{\zeta^2}{c_s^2}Q + \frac{\zeta^2k_x^2}{D}W \notag\\ 
%   &=\left(\frac{\ii
%       k_x r}{D}\frac{d\Omega^2}{dz} -
%     \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
%     \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
%   &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
%     \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
%   &\zeta^2W - \frac{\gamma}{\Gamma}\zeta^2 Q +
%   \frac{\ii\zeta}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
%   &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
%   \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
% \end{align}
% where
% \begin{align}
%   D \equiv \kappa^2 - \zeta^2,
% \end{align} 
% and we have replaced the notation $\p_z$ by $d/dz$ since we are
% considering solutions at a fixed radius.

% Extrapolating results obtained from the VGSB to the global disk has
% limitations. This is because within the VGSB framework, $d\Omega/dz$
% can be freely specified. 
% We choose $d\Omega/dz$ to be that due to the thermal structure in 
% the global disk (Eq. \ref{vertical_shear}). However, background radial
% gradient terms are ignored elsewhere in the VGSB (which is done by
% setting $\hat{g}_c=0$).

% We further discuss caveats of the VGSB in Appendix
% \ref{global_corr}. We find that for fast thermal relaxation
% ($\beta\ll1$) and perturbations with small radial
% lengthscales ($k_xH\gg 1$), which characterizes the VSI, the VGSB
% framework is adequate for the purpose of a simple analytical
% treatment.