### 2015-04-25 paper.tex
edits mainly to section 2.  Please proofread and check.  It should be obvious but I mainly wanted to (aO cut ideas already in the intro (2) expand on the Richardson like argument (i.e not force the reader to go to Urpin to follow the flow) (c) put the figs a bit earlier (e.g. fig 1 where we first give shear and buoyancy) and (d) expand a bit the physical discussion of t_c figure and implications, but I could be convinced that some of this belongs more in the concluding discussion (on this is was just worried, maybe too much, about giving some wrong impressions).

%\documentclass[12pt]{aastex}
\documentclass[iop]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
\usepackage{paralist}
\shorttitle{Vertical shear instability}
\shortauthors{M-K. Lin, A. Youdin}

\usepackage{amsmath}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\zmax}{z_\mathrm{max}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\khat}{\hat{k}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\w}{ \widetilde{W}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\actaa}{Acta Astron.}
\defcitealias{lin12}{L12} 
\defcitealias{nelson13}{N13}
\defcitealias{barker15}{BL15}
\defcitealias{mcnally14}{MP14}
\def \OmK {\Omega_{\rm K}}

\begin{document}

\title{Cooling Requirements for the Vertical Shear Instability 
  in Protoplanetary Disks}
%\title{Vertical shear instability in astrophysical disks: vertically
%global linear calculations with thermal relaxation and applications
%to the Solar Nebula}  
\author{Min-Kai Lin \& Andrew Youdin}
\affil{Department of Astronomy and Steward Observatory, \\
University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
\email{minkailin@email.arizona.edu, youdin@email.arizona.edu}

\begin{abstract}
  It is difficult to understand how cold circumstellar disks accrete onto their central stars.  Due to 
  lower levels of ionization, the magnetic transport mechanisms that
  operate in warmer disks are less effective.  
  A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
  transport in cold accretion disks such as protoplanetary disks (PPDs).  
  The VSI is driven by a weak vertical gradient in the disk's orbital motion.
  In order to grow, the VSI must overcome vertical buoyancy, a strong
  stabilizing influence in cold disks, where  
  heating is dominated by external irradiation.  Rapid cooling, via
  radiative losses, reduces the effective buoyancy 
  and allows the VSI to operate.  In this paper, we quantify the
  cooling timescale, $t_c$, needed for growth of the VSI.
   We perform a linear analysis of the VSI with cooling in
  vertically global and radially local disk models. 
  For irradiated disks, we find that the VSI is most vigorous for
  rapid cooling with $t_c < \OmK^{-1}  h |q| / (\gamma -1)$ 
  in terms of the Keplerian orbital frequency, $\OmK$; the disk's
  aspect ratio, $ h \ll 1$; the radial  
  powerlaw temperature gradient, $q$; and the adiabatic index,
  $\gamma$.  For longer cooling times, the VSI  
  is much less effective because growth slows and shifts to smaller length scales, which are more prone to 
  viscous or turbulent decay.  We apply our results to PPD models 
  where $t_c$ is determined by the opacity of dust grains.  We
  find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a
  characteristic growth time of $\sim 30$ local orbital periods.
  Growth is suppressed by long cooling times both in the opaque  
  inner disk and the optically thin outer disk.  A reduction in the
  dust opacity by a factor of 10 increases cooling times  
  enough to quench the VSI at all disk radii.  Thus the formation of
  solid protoplanets, a sink for dust grains, can impede the VSI. 
\end{abstract}

\section{Introduction}\label{intro}
Understanding how disks transport mass and angular momentum underlies 
many problems in astrophysics, including star and planet formation  
\citep[][]{lyndenbell74, armitage10}.  The turbulence associated with many transport mechanisms
is particularly important for dust evolution and planetesimal formation \citep{yl07, chiang10}.

Magneto-hydrodynamic (MHD) turbulence driven by
the magneto-rotational instability \citep[MRI,][]{balbus91} has long been the most
promising transport mechanism in low mass disks with weak self-gravity. 
However, many parts of protoplanetary  
disks (PPDs) are cold, have low levels of ionization, and do not support the MRI 
\citep{blaes94,salmeron03}. Recent simulations
suggest that significant portions of PPDs fail to develop MHD
turbulence \citep[e.g.][]{simon13, lesur14,bai15,gressel15}.

A purely hydrodynamic mechanism could circumvent difficulties with non-ideal MHD, but must overcome
the strong centrifugal stability imposed by the positive radial specific angular
momentum gradient in nearly Keplerian disks \citep{balbus96}.
One possible route to hydrodynamic turbulence is the vertical shear
instability (VSI, \citealp{urpin98, urpin03, nelson13}, hereafter \citetalias{nelson13}).  The
basic mechanism of the VSI  in disks was first 
identified in the context of differentially rotating stars \citep[][]{goldreich67,fricke68}.   
The VSI arises when vertical shear, i.e. \ a variation in orbital motion along the rotation axis, 
destabilizes inertial-gravity waves, which are oscillations 
with rotation and buoyancy as restoring forces. Vertical shear
occurs wherever the disk is baroclinic, i.e.\ when constant 
density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
practically unavoidable in astrophysical disks, except at special locations like the
midplane.

To overcome centrifugal stabilization, the VSI triggers motions which are vertically elongated 
and radially narrow.  %This geometry minimizes the effects of the radial gradient of angular momentum.
Vertical elongation taps the free energy of the vertical shear \citep{umurhan13}, but is also subject to 
the stabilizing effects of vertical buoyancy if the disk is stably
stratified. To overcome vertical buoyancy, the VSI requires a short cooling time
(\citealp{goldreich67}; \citetalias{nelson13}). Rapid radiative 
cooling, equivalently a short thermal relaxation timescale, brings a
moving fluid element into thermal equilibrium with 
its surroundings, thereby diminishing buoyancy. %not obvious actually
                                %- requires assumption of ideal gas
                                %e.o.s. and pressure eqm established
                                %quickly 
An isothermal equation of state implies instantaneous cooling (and
heating), and is the ideal context for studying the VSI
(\citealp{urpin03}; \citetalias{nelson13}; \citealp{mcnally14}, hereafter \citetalias{mcnally14}; \citealp{barker15},
hereafter \citetalias{barker15}).

Alternatively, vertical buoyancy can be eliminated by strong internal
heating, i.e.\ by the onset of convection, so that the disk becomes 
neutrally stratified in the vertical direction
\citepalias{nelson13,barker15}. However, realistic PPDs should be vertically
stably stratified in the outer regions, beyond $\sim 1$--5 AU, where
heating is  irradiation dominated \citep{bitsch15}.  In the inner
disk, strong vertical buoyancy is possible if accretion heating is
weak or is concentrated in surface layers \citep{gammie96,lesniak11}.

A full understanding of the VSI in real disks therefore requires
considering finite, non-zero cooling times, $t_{\rm c}$. Non-linear
hydrodynamical simulations with a prescribed  $t_{\rm c}$ find that VSI
turbulence in stably stratified disks requires rapid cooling  
with $t_{\rm c} \lesssim 0.01P_{\mathrm{orb}}$, where $P_\mathrm{orb}$ 
%=2\pi/\Omega$ 
is the local orbital period %and $\Omega$ is the orbital
%frequency 
 \citepalias{nelson13}. \citetalias{nelson13} further find
that the VSI can drive moderately strong transport, with Reynolds
stresses of $\alpha \sim 10^{-3}$ times the mean pressure. Simulations
with realistic radiative transfer, in lieu of a parameterized $t_{\rm
  c}$, find that the VSI in irradiated disks drive transport with
$\alpha \sim 10^{-4}$  in a  $\sim 2$ -- 10AU PPD model \citep{stoll14}.

To reach hydrodynamic turbulence via the VSI, a disk should  
first develop the linear instability. A quantitative 
examination of the thermodynamic requirement for the linear VSI is
thus essential in assessing its 
relevance to PPDs. In this work, we present a
comprehensive study of the linear VSI in vertically global disk models 
with non-zero cooling timescales, which permit buoyancy forces. 
% In this work, we present a comprehensive study of the linear VSI in vertically global disk models 
% with non-zero cooling timescales $t_c > 0$, which permit buoyancy forces. 
% Our effort generalizes both early analyses of the VSI in vertically
% localized disk models \citep{urpin98,urpin03} and recent calculations in vertically global 
% disks \citepalias{nelson13,mcnally14,barker15}. These works either
% consider the limit of instantaneous cooling ($t_c\to0$) or neutrally
% stratified disks, for which buoyancy forces play no role. 
Our effort generalizes previous studies of the linear VSI in 
vertically local \citep{urpin98,urpin03} and vertically global 
\citepalias{nelson13,mcnally14,barker15} disk models that consider the limit
of instantaneous cooling, for which buoyancy plays no role.  
%to study of how finite cooling times, and thus buoyancy, 
%affect the VSI. Our effort may also be regarded
%as a extension of early analyses of the VSI in 
%vertically localized disk models with cooling \citep{urpin98,urpin03} to
%vertically global disks with cooling. 
We consider disk  
models with vertically constant $t_{\rm c}$, as well as more realistic
cooling timescales based on dust opacity in both the optically thick
(diffusive) and thin regimes, where appropriate. Coupled with the 
results from numerical simulations, our results can be used to 
understand how the VSI operates over a wide range of model PPDs.

%\emph{TODO: the following is still a bit too wordy IMO but we can redo as a final step after these sections are finalized.}

This paper is organized as follows. In \S\ref{vsi_require} we 
introduce and qualitatively explain our main result of a critical or
characteristic cooling timescale, beyond which the VSI is either
suppressed or inefficient. We then present the  
formal linear problem in \S\ref{setup}---\ref{numerical}.    
In \S\ref{setup} we list the governing equations and global
equilibrium disk models under consideration, and formulate the linear
stability problem in \S\ref{linear}. In 
\S\ref{analytical} we give a simplified analytical discussion for
vertically isothermal disks, where we derive the aforementioned
characteristic upper limit on the cooling or thermal timescale for the VSI. 
% Our treatment is grossly simplified owing to a
% number of approximations made, but closely parallels previous work, 
% allowing us to highlight the essential effects of vertical shear, and to
% derive a characteristic upper limit to the thermal timescale for the
% VSI. This timescale is \emph{much shorter} than the dynamical
% timescale. 
In \S\ref{numerical} we numerically
solve the linear problem, and  
find basic agreement with our analysis.
% , including
% the aforementioned characteristic thermal timescale. 
In 
\S\ref{application} we apply our results to the Minimum Mass
Solar Nebula (MMSN) by modeling cooling timescales based on realistic
disk properties, and find that the VSI can operate effectively in the
outer parts of the MMSN. We summarize in \S\ref{summary} with a
discussion of caveats and possible directions for future work.

%The VSI is an axisymmetric linear instability 
%in three-dimensional (3D) disks in which the rotation frequency
%$\Omega$ depends on the height $z$ from the disk midplane. 
%Vertical shear occurs whenever the disk is baroclinic, i.e.\ when constant 
%density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
%practically unavoidable in disks, except at special locations like the midplane.
%
%A height-dependence in $\Omega$, or vertical shear, can render
%inertial waves unstable. \cite{barker15} gives a simple 
%interpretation of the VSI: in an axisymmetric, inviscid, neutrally
%stratified disk, the exchange of fluid parcels vertically, while
%conserving their specific angular momentum, can result in a new
%configuration of lower energy if the distribution of specific angular
%momentum varies with $z$. The VSI is characterized by perturbations
%with large vertical length-scales  relative to radial length-scales.   
%
%Early analysis of the VSI in radially and vertically local 
%models of accretion disks \citep{urpin98,urpin03} showed that the VSI
%requires a short thermal timescale so that perturbations evolve 
%isothermally, unless the disk is neutrally stratified in the vertical
%direction. This is because
%vertical motions associated with the VSI 
%are stabilized by buoyancy forces if the disk is sub-adiabatically
%stratified (convectively stable). A sufficiently short thermal
%timescale renders buoyancy ineffective, but a quantitative criteria is
%desirable. This is our primary goal.  
%
%The need for a short thermal timescale for the VSI to operate in
%stably stratified disks has been demonstrated in recent numerical simulations 
%\citep{nelson13,stoll14}. When the  thermal timescale $t_c$ is
%parametrized in units of the orbital period  $P_\mathrm{orb}$,
%\cite{nelson13} find that $t_c\lesssim  0.01P_\mathrm{orb}$ was needed
%to observe the VSI. That is, the  disk temperature needs to remain
%effectively unchanged. Radiation hydrodynamic simulations performed by
%\cite{stoll14} also showed that the VSI is only sustained with the
%inclusion of heating from stellar irradiation, which results in an
%effective thermal timescale $t_c\ll P_\mathrm{orb}$ that maintains a
%fixed temperature profile.     
%
%These studies show that the VSI is generally sensitive to disk
%thermodynamics. It is thus important to quantify this dependence, for
%example on disk parameters, in order to evaluate the relevance of the
%VSI to realistic disks, which one may expect to be stably
%stratified and have non-zero thermal timescales.  In order to determine
%where in the disk the VSI may operate, and hence lead to hydrodynamic
%turbulence, understanding the thermodynamic dependence of the linear VSI is essential. 
%
%In this work we study the linear VSI in radially local, vertically
%global models of astrophysical disks. This approach has
%been adopted in recent studies of the linear VSI
%\citep{nelson13,mcnally14,barker15}. However, these studies consider 
%isothermal perturbations, for which the thermal timescale is formally
%zero, and/or neutrally stratified disks in which buoyancy forces are
%absent by construction. We account for both effects to study the
%VSI in more general disk models. To do so, we examine 
%sub-adiabatically stratified disks and include an energy equation 
%with a source term that allows the thermodynamic response of 
%perturbations to be varied smoothly between isothermal and 
%adiabatic.   
%
%
%This paper is organized as follows. We first highlight in
%\S\ref{vsi_require} the need of a very short thermal timescale for the
%VSI to operate in stably stratified disks. There we also preview our
%main result --- a quantitative thermodynamic criteria for the VSI ---
%and a first application to protoplanetary disks. The remaining sections
%are devoted to deriving these results from linear theory. 
%
%In \S\ref{setup} we list the governing equations and global
%equilibrium disk models under   
%consideration. We formulate the linear stability problem in 
%\S\ref{linear} and simplify it by making the radially local
%approximation. We study the linear problem using two approaches. In
%\S\ref{analytical} we present an analytical discussion of vertically
%isothermal disks. Our treatment is grossly simplified owing to a
%number of approximations made, but closely parallels previous work, 
%allowing us to highlight the essential effects of vertical shear, and to
%derive a characteristic upper limit to the thermal timescale for the
%VSI. This timescale is \emph{much shorter} than the dynamical
%timescale. In \S\ref{numerical} we numerically
%solve the linear problem in the radially local approximation, but
%relax all other approximations made in our analytical approach.    
%We find basic agreement with our analysis, including
%the aforementioned characteristic thermal timescale. 
%
%In
%\S\ref{application} we apply our results to the Minimum Mass
%Solar Nebula (MMSN) by modeling thermal 
%timescales based on realistic disk properties, and find that the VSI
%can operate effectively in the outer parts of the MMSN. We summarize in
%\S\ref{summary} with a discussion of caveats and possible directions
%for future work.

\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}     
In a stably stratified thin disk, the VSI  
requires a thermal timescale $t_c$ much shorter than the
disk dynamical timescale, 
\begin{align}
  t _c \ll \OmK^{-1}. 
\end{align}
where $\OmK$ is the Keplerian frequency (N13).    
This requirement -- that the cooling time be much shorter than an orbital period, 
which in turn is much smaller than the relevant oscillation period of vertically elongated gravity waves  -- is quite stringent.
It highlights the fact that vertical buoyancy is a very strong stabilizing influence.
Very strong radiative damping is thus required to weaken buoyancy and allow the weak vertical shear to drive instability.

To roughly quantify the required smallness of $t_c$, we consider a
vertically isothermal disk with aspect-ratio $ h$, 
 radial temperature profile $T \propto r^q$ and adiabatic index
$\gamma$. For a PPD, $ h \sim 0.05$, $q\sim -0.5$ and $\gamma\simeq 1.4$.

\begin{figure*}
  \includegraphics[width=\linewidth]{figures/rhoshearNz}
  \caption{The radial and vertical structure of a ``minimum mass" PPD model showing gas density (left, relative to the midplane density at 1 AU), vertical shear rate (middle) and
    buoyancy frequency (right).
    \label{eqm_structure} 
  }
\end{figure*}

For a thin disk, the vertical shear increases with height, $z$, from the midplane as 
\begin{align}\label{vshear_thin}
  r \frac{\p \Omega }{\p z} \simeq  h q \OmK \left(\frac{z}{2 H}\right),
\end{align}
for several disk scale heights, $H = hr$. 
%As noted in
%\S\ref{intro} this vertical variation is a source of free energy and
%thus vertical shear ($\p_z\Omega \neq0$) is destabilizing. 
%
%For $\gamma>1$ the disk is stably stratified with a real 
This destabilizing shear completes with the stabilizing vertical
buoyancy frequency $N_z$.  In a thin disk,     
\begin{align}\label{nz_thin}
  N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \left(\frac{z}{H}\right)^2
  \OmK^2.  
\end{align}
Fig. \ref{eqm_structure} maps shear and buoyancy (as well as the gas density) 
in a fiducial PPD model, whose parameters are in Appendix \ref{mmsn} . %Here $\rho_0$
%is the midplane density and vertical shear rate and buoyancy
%frequencies are normalized by the local Keplerian frequency. 
Vertical shear is 
generally weak compared to buoyancy, suggesting stability.

Without any cooling, the Solberg-Hoiland criteria confirms that vertical buoyancy is strong
enough to ensure (axisymmetric) stability, see \S\ref{solberg}.  With radiative cooling, 
thermal fluctuations decay which, combined with pressure equilibration, reduces the effective buoyancy.

%However, the stabilizing effect of buoyancy can be mitigated for 
%effectively isothermal disturbances. Consider a fluid element displaced upwards, away from
%the disk midplane. In a stably stratified adiabatic disk, it would
%sink because of buoyancy. However, if thermal timescales are short, in
%that temperature flucations decay rapidly, then the fluid element would attain its surrounding 
%density and  
%%(assuming an ideal gas and it reaches pressure 
%%equilibrium with its surroundings). 
%thus buoyancy is rendered ineffective. Qualitatively, a short thermal
%timescale $t_c$ reduces the effective (stabilizing) buoyancy force.

How short must $t_c$ be for vertical shear to prevail?  We start by 
considering the Richardson number $\mathrm{Ri} = N_z^2/(r\p_z\Omega)^2$, a 
ratio of buoyant to shear energies.  Though not precisely applicable to our problem,
non-rotating shear flows are stable if $\mathrm{Ri} > 1/4$ (\citealp{chandrasekhar61}, 
also see \citealp{ys02, lee10} for applications to thin dust layers in PPDs).
Following \cite{urpin03} and \cite{townsend58},  we reduce the buoyant energy by the 
ratio of cooling to forcing timescales, $t_c |r \p_z\Omega| \leq 1$ (with no reduction in buoyant energy 
expected when this inequality fails).  With this reduction, the Richardson like criterion for 
shear \emph{instability} becomes
\begin{align}\label{bcrit_est}
  t_c \lesssim \frac{\left|r\p\Omega/\p z\right|} {N_z^2}. 
\end{align}

This result is far from rigorous, but it contains some useful insights.
 If we interpret Eq.\ \ref{bcrit_est} as a local criterion, then we see that 
 the minimum cooling time decreases with height as $t_c \propto 1/z$.
 This finding is relevant to the radiative damping of so called ``surface 
 modes," which we consider in \S\ref{numerical}.

 To roughly obtain a global cooling time criterion, we evaluate Eq.\ \ref{bcrit_est}
 at $z \sim H$.  To get the precise form we want, we choose $z = \gamma H / 2$ to 
 obtain the cooling requirement
 \begin{align}\label{prelim_bcrit}
 t_c <  h |q| \OmK^{-1}/ (\gamma - 1)
\end{align}   
for the VSI in a vertically isothermal, thin disk.  This estimate agrees with 
our central result, the cooling criterion that we derive more rigorously in \S\ref{iso_vsi_beta_crit}.

%
%The left-hand side represents an effective buoyancy, which is
%stabilizing; while the right-hand side represents vertical shear,
%which is destabilizing. In a thin disk $ h \ll 1$, so the destabilizing effect of 
%vertical shear is rather weak because $|r\p_z\Omega|\sim h\OmK
%\ll \OmK$. By comparison, the buoyancy frequency $N_z\sim
%\OmK$. It follows that $t_c\lesssim O( h\OmK^{-1})$ is
%required for instability.  \emph{In a thin  
%  disk the thermal timescale must be shorter than the dynamical
%  timescale by $O( h)\ll 1$ for the VSI to operate.}   
%
%
%Then even a small (positive)
%vertical entropy gradient can significantly stabilize the VSI. 
%Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
%with height. Eq. \ref{bcrit_est} then suggest that with
%increasing height, buoyancy would eventually stabilize the VSI (since
%$N_z^2\propto z^2$ and $\p_z\Omega\propto z$).
%% This means that 
%% disturbances localized away from the disk midplane are easily stabilized by
%% buoyancy. 
%This argument assumes vertically localized disturbances. It is,
%however, not clear how to apply Eq. \ref{bcrit_est} to vertically
%global disturbances, as considered in this work.
%
%
%
%
%Our central result is the requirement 
%\begin{align}\label{prelim_bcrit}
%  (\gamma - 1)t_c <  h |q| \OmK^{-1}
%\end{align}   
%for the VSI in a vertically isothermal, thin disk. This equation
%reflects Eq. \ref{bcrit_est} but there is no $z$ dependence. 
%% While assumptions and simplifications will be made 
%% to derive this, 
%Eq. \ref{prelim_bcrit} can be 
%readily applied to evaluate where in a disk the VSI is likely to
%operate.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.7cm 0cm
  0.73cm]{figures/bcrit_mmsn2d_kx10}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.46cm 0cm
  0.73cm]{figures/bcrit_mmsn2d_kx50}
  \caption{Estimate of the thermal timescale  in a fiducial 
    PPD model, the Minimum Mass Solar Nebula. The red
    line corresponds to a characteristic thermal timescale, derived in
    this work, below which the VSI is expected to operate effectively
    (the region outside the line).    
    \label{bcrit_mmsn2d} 
  }
\end{figure}

We now consider whether, and where, thermal timescales in realistic PPDs
are short enough to support VSI growth.  Fig. \ref{bcrit_mmsn2d} plots the
 thermal timescales in our fiducial PPD model with a small grain dust opacity. 
 Since cooling by radiative diffusion depends on a perturbation's lengthscale, we
 plot $t_c$ for two values of the radial wavenumber,  $k_x$,  the relevant 
 component for VSI modes that are radially narrow.   The red line
marks the boundary implied by Eq. \ref{prelim_bcrit}.  We caution that 
 Eq. \ref{prelim_bcrit} applies when $t_c$ is vertically uniform.  This section is 
 aimed at basic insights, while \S\ref{application}
 analyzes the VSI with realistic spatial variations in $t_c$.

Due to the smaller values of $t_c$ in the outer disk, we generally expect the VSI to 
operate outside $\sim 10$ AU.  This conclusion changes however, if the dust opacity is reduced, 
slowing down cooling, as we show later (e.g.\ Fig.\ \ref{mmsn_bcrit_bcool}).

In the inner disk we see that thermal times are long in the midplane, but short at higher altitutdes. 
Again, this structure necessitates the more detailed analyses that follow.  Nevertheless, we can already 
see that due to radiative diffusion higher wavenumbers can cool more rapidly.  Thus smaller scale disturbances 
may trigger VSI at smaller disk radii.

However we cannot and do not find that VSI will 
always be relevant for sufficiently small perturbations.  For starters, very small wavelengths are in the optically 
thin cooling regime,  where cooling is independent of lengthscale.  Moreover, small scale perturbations 
are prone to viscous damping, and should drive weaker transport.

We now proceed with our detailed analysis of the VSI to arrive
at Eq. \ref{prelim_bcrit} and verify it numerically. The interested 
reader may jump to \S\ref{application} for the application of our
results to PPDs.

\input{setup}
\input{linear_iso}
\input{linear_adia}
\input{numerical}
\input{application} 
\input{summary}

\appendix
\input{appendix}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}