### 2015-05-05 application.tex
Comments on Section 7 for suggested edits.  Hopefully they are coherent, but ask if not.

\section{Application to protoplanetary disks}\label{application} 
\emph{For this section. There are 3 main ideas, and I think they should be the subsections, more comments related to this are in the text itself:
\begin{itemize}
\item Developing the cooling model in opt thin and this regimes, including MMSN value
\item Comparison to expected $\beta_{crit}$
\item Linear growth with MMSN cooling law
\end{itemize}}

\emph{Suggested edits for this intro
\begin{itemize}
\item Instead of starting over again with the full energy equation focus on giving a phyiscal values for $\Lambda$ and $t_c$ already introduced.
\item Don't describe what we don't do, especially at the outset.  Say what's done, and then if essential list caveats.
\end{itemize}}

As an application of our linear models, we estimate where in a
PPD do the thermodynamic conditions allow the VSI to
operate. For a realistic disk one might consider an energy equation
with radiative cooling, 
\begin{align}\label{real_energy}
  % \rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
  \frac{\p P}{\p t} + \bm{v}\cdot\nabla P +\gamma P \nabla\cdot\bm{v} = - \frac{P}{\rho T
    C_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
\end{align}
where $C_v$ is the heat capacity at constant
volume, 
$\bm{F}$ is the radiative flux and $Q_+$ is a heating
term to permit an equilibrium solution.

One should now model in detail each of the terms on the RHS of
Eq. \ref{real_energy}, linearize them, then solve 
the resulting linear eigenvalue problem. This is beyond 
the scope of the present work. Here, we take the simpler approach of
representing the linearized RHS of Eq. \ref{real_energy} by a thermal
relaxation time $t_c$. To do so, we first consider separately
perturbations with radial lengthscales $l\gg l_\mathrm{rad}$ and  
$l\ll l_\mathrm{rad}$, where      
\begin{align}\label{lrad}
  l_\mathrm{rad} \equiv \frac{1}{\kappa_d\rho} 
\end{align} 
is the photon mean-free-path and $\kappa_d$ is the (dust) opacity.

\subsection{Radiative diffusion}
\emph{I think 7.1-7.4 should be combined to a single section on the "Cooling Model".  As that clearly is a single idea and osme of the subsections are too trivial. Conversion to subsubsections as needed. However the application to beta-crit could be split off as 7.2.  Otherwise that material which has a crucial figure discussion is a bit buried.}

\emph{the following sentence is too long and has a dangling "this".}

For perturbations with lengthscales $l\gg
l_\mathrm{rad}$, we assume temperature fluctuations are smoothed out
by radiative diffusion and model this in linear theory by writing the
linearized RHS of Eq. \ref{real_energy} as

\emph{I would have this be an equation for something, like $\Gamma$ or $t_c$.}
\begin{align}\label{diff_cool_proper}
  \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
    T\right),  
\end{align}
where $k_\mathrm{rad}$ is the conduction coefficient to be defined
implicitly.

\emph{To get to the point more directly and intuitively, I would start below with "We include only the radial derivatives in Eq.\ blah, neglecting vertical derivatives.  This approximation is valid for the vertically elongated modes we consider."  I think we've already discussed vertical elongation enough that this can be brief.  Then maybe say that keeping the vertical derivates would be bad, if you really want to.  "With this simplification (give revised equation)}

However, Eq. \ref{diff_cool_proper} fundamentally changes the linear
problem by introducing higher order derivatives. We defer a proper
analysis of this problem to a follow-up study. Here, we are interested
in order-of-magnitude estimates for perturbations with radial
lengthscales much smaller than vertical lengthscales. We thus proceed 
by assuming vertical derivatives can be neglected in 
Eq. \ref{diff_cool_proper} and write  
\begin{align}\label{diff_cool_approx}
  \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
    T\right) &\to -\frac{k_\mathrm{rad}}{\rho
    C_v}k_x^2P\left(\frac{\delta T}{T}\right)\notag\\
  &\equiv -\eta k_x^2 \left(\delta P - \frac{P}{\rho}\delta\rho\right), 
\end{align}
\citepalias[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho C_v$ is
the diffusion coefficient and 
is given in terms of physical disk parameters as 
\begin{align}\label{eta_def}
  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 C_v}, 
\end{align}
where $\sigma_s$ is the Stefan-Boltzmann constant. 
From Eq. \ref{diff_cool_approx} we identify the scale-dependent thermal relaxation
time for diffusion as 
\begin{align}\label{tc_diff_cool} 
  t_\mathrm{diff} = \frac{1}{\eta k_x^2}.%\equiv \frac{\OmK^{-1}}{\hat{\eta}\khat^2}, 
\end{align}

\subsection{Newtonian cooling}\label{newton_cool}
\emph{There are no rules against subsections this short, but I think it's bad style.}

For perturbations with small lengthscales, $l\ll l_\mathrm{rad}$, 
the disk is optically thin.\\
\emph{Above, the disk isn't (necessarily) opt. thin, the pert. is} \\
 In this case, we assume 
Newtonian cooling can be applied, which is the thermal
relaxation model as adopted in our linear models. That is, temperature
fluctuations $\delta T$ decay on a timescale $t_\mathrm{cool}$,
independent of $k_x$. For simplicity, we take  
\begin{align}
  t_\mathrm{cool} = \frac{l_\mathrm{rad}^2}{3\eta}. 
\end{align}
We note that $t_\mathrm{cool}$ does not explicitly depend on $\rho$.

\subsection{A model for combined thermal relaxation}\label{toy_relax}
%To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, 
We define the effective thermal relaxation timescale in linear theory as
\begin{align}\label{tc_def}
  t_c &\equiv t_\mathrm{cool} + t _\mathrm{diff}. %=
  % \frac{l_\mathrm{rad}^2}{3\eta} + \frac{1}{\eta k_x^2},  
\end{align}
Eq. \ref{tc_def} is a simple prescription so
that for small scales (large $k_x$), $t_c\to t_\mathrm{cool}$, while
for large scales (small $k_x$), $t_c\to t_\mathrm{diff}$. Writing the
volume density $\rho = \Sigma\hat{\rho}/\sqrt{2\pi}H$, where
$\hat{\rho} = \rho/\rho_0$ and $\Sigma(r)$ is the surface density at the
radius of interest, the dimensionless thermal
relaxation time $\beta$ becomes 
\begin{align}\label{real_beta}
  \beta(r,z;\khat) \equiv t_c\OmK =
  % \frac{1}{\hat{\eta}}\left[\frac{l_\mathrm{rad}^2}{3H^2}
  %   + \frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right]  \simeq
  \frac{\Sigma^2\OmK}{\eta\rho^2}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
    + \frac{\hat{\rho}^2(z)}{2\pi \khat^2}\right].
\end{align}
Note that $\beta$ depends the perturbation lengthscale $\khat$,
distance from the midplane, as well as the distance from the star.

\subsection{Evaluation for a fiducial disk model}
We evaluate Eq. \ref{real_beta} for the PPD model
described in \cite{chiang10} and summarized in Appendix \ref{mmsn}
(where an opacity model is also chosen). Defining $r_\mathrm{AU}\equiv 
r/\mathrm{AU}$, we find Eq. \ref{real_beta} becomes
\begin{align}\label{beta_mmsn}
  \beta(r, z; \khat) =
  &\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\notag\\ 
&\times\left[\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}r_\mathrm{AU}^{33/7}+\frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right].         
\end{align} 
We refer to the case with $\gamma=1.4$, $\mu=2.33$ and the scalings
$\hat{\Sigma}=\hat{\kappa}_d=\hat{T}=1$ as the Minimum Mass Solar
Nebulae (MMSN) and $\beta=\beta_\mathrm{MMSN}$, \\
\emph{I would keep $\kappa$ scaling in this eqn, since it's used in plots.  Then just say  that $\kappa_d = 1$ is fiducial case. }
\begin{align}\label{beta_mmsn_simp}
  \beta_\mathrm{MMSN} = 3.9\times10^{-3} r_\mathrm{AU}^{9/14}\left( 1 +
    1.9\times10^7\hat{\rho}^2r_\mathrm{AU}^{-33/7}\khat^{-2}\right). 
\end{align}
\emph{I would explain here that when $k < c\rho \kappa^m r^n$ (with correct values for $c, m, n$) cooling is optically thin, as this is useful for explaining figure.}
% where
% \begin{align}
%   \
% %  \frac{3.7\times10^{-3}}{\mu\left(\gamma-1\right)\hat{\kappa}_d\hat{T}^5}, 
% \end{align}
% and
% \begin{align}
%   \khat_0 = 4.4\times10^{3}\hat{\kappa}_d\hat{T}^2\hat{\Sigma}.
% \end{align}

\emph{Sub-section break  ( 7.2) for the "$\beta_{crit}$ in the MMSN" as this is break from the derivation (and evaluation) of the cooling time (which is fairly general) to the application to the VSI.}

It is instructive (\emph{advice: don't say that "it is instructive", just be instructive}) to compare the above thermal timescale 
with the criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
\begin{align}\label{bcrit_mmsn}
  % \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
  % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
  \beta_\mathrm{crit,MMSN} = 0.024r_\mathrm{AU}^{2/7}. 
\end{align}

\emph{This paragraph currently merely copies the figure caption, which is a waste and not illuminating to read.  Better to give some description of what the figure is showing. (see below for ideas)}\\
In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
=2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for opacity scales
$\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. \\
\emph{Here you can describe the general features e.g. diffusive and thin cooling regions, which will help later discussions.}\\
We also plot Eq. \ref{bcrit_mmsn} in each panel.  \emph{This sentence explains nothing and is likely to be of little help to the reader.  Instead wait to describe the science, e.g. after describing the general shape of the $\beta$ curves,  you can start the discussion below with "Fig () also compares the MMSN cooling times to $\beta_{crit}$.}

Consider first the MMSN shown in the middle panels of 
Fig. \ref{mmsn_bcrit_bcool}. The critical thermal timescale,
Eq. \ref{bcrit_mmsn}, is useful in the optically thin regime (where the
curves overlap), since in this limit $\beta$ does not depend on
height, as assumed in our analysis. \\
\emph{above sentence too long and confusing.  "useful" is a vague term.  
"where the curves overlap" is also confusing, forcing the reader to look at the figure in the middle of a complex sentence.  Better to have already explained above how the $\beta$ curves show that opt. thin cooling is height indep. (even though should be obvious from equations).  Then the comparison to $\beta_{crit}$ can be made as something like "Since beta-crit was derived for a height-independent cooling time,  the direct comparison our MMSN cooling model is valid only for modes with optically thin cooling. "} 
We thus infer that the VSI can
operate at tens of AU in the MMSN.

In the inner few AU of the MMSN, however, $\beta$ 
may be smaller or greater than $\beta_\mathrm{crit}$ depending on
$z$ as the disk transitions from an optically thick midplane to an
optically thin atmosphere (see also Fig. \ref{bcrit_mmsn2d}). 
%Given a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at
%the optically thick midplane, but with increasing height the disk 
%becomes optically thin and $\beta<\beta_\mathrm{crit}$. 
Our analytical criteria $\beta < \beta_\mathrm{crit}$ is not valid for
vertically non-uniform $\beta$, and we must turn to numerical
solutions of the linear problem to determine whether or not the VSI
operates. We do this in the next section.    \\
\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}

Lowering the opacity, as shown in the top panels of
Fig. \ref{mmsn_bcrit_bcool}, shifts the curves inwards but also raises
them.\\
\emph{this graphical discussion is not very inuitive (the main problem) and even arbitrary (lowering vs. shifting to the left).  Better to say what's happening physically, i.e. `With higher opacity, optically thin cooling faster and diffusive cooling is slower, explaining how the curves shift.} \\
In fact, lowering the opacity by an order of magnitude renders  
$\beta > \beta_\mathrm{crit}$ even away from the midplane. This is 
expected to stabilize the disk. Raising the opacity, as shown in 
the lower panels of Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal
timescales, but shifts the curves outward, so the instability is
restricted to larger radii.

\begin{figure*}
  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
  0cm]{figures/bcrit_mmsn_kap0d1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 1.8cm 0cm
  0cm]{figures/bcrit_mmsn_kap0d1_z3}\\ 
  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
  1cm]{figures/bcrit_mmsn_kap1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm
  1.8cm 0cm 1cm]{figures/bcrit_mmsn_kap1_z3}\\
  \includegraphics[scale=.47,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/bcrit_mmsn_kap10_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 0cm 0cm
  1cm]{figures/bcrit_mmsn_kap10_z3} 
  \caption{Dimensionless thermal relaxation timescales $\beta$,
    evaluated at the midplane (left) and at $z=3H$ (right) in the
    fiducial PPD. Eq. \ref{beta_mmsn} is plotted  
    for three values of the 
    perturbation radial wavenumber: $\khat=1$ (dotted), $\khat=10$
    (dashed) and $\khat=100$ (dashed-dot), for three values of the
    opacity relative to the MMSN: $\hat{\kappa}_d=0.1$ (top),
    $\hat{\kappa}_d=1$ (middle) and $\hat{\kappa}_d=10$ (bottom).  
    The solid line is the 
    critical thermal relaxation timescale $\beta_\mathrm{crit}$.  
    \label{mmsn_bcrit_bcool}}   
\end{figure*}

\emph{the following statement is not generally true, so it should not be stated without caveat}

We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$ 
for sufficiently large $\khat$. We illustrate this in 
Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at each radius the value
of $\khat$ such that $\beta = \beta_\mathrm{crit}$. If we consider the
midplane of the inner disk, say $r\sim 1$AU, the VSI would only occur
on very small radial lengthscales $\ll O(10^{-2}H)$. In the disk
atmosphere where $\beta$ is small, surface modes may develop, 
%However, at such
%small scales, surface modes are likely to dominate (and are
%restricted to the disk atmosphere), 
but they are not well-defined in vertically isothermal disk models
with an imposed boundary, and they also require large $\khat$.

%surface modes can develop for large $\khat$, 
%which have small vertical lengthscales scales. 
%also small vertical scales if they are surface modes

\emph{The following viscosity discussion might be better placed elsewhere.  Since the issue also comes up for numerical solutions (not just the comparison to beta-crit here), maybe a short concluding subsection to summarize all the arguments on the "vialibity of short wavelength VSI".}

However, disturbances with large $\khat$ may be subject to viscous
decay in a real disk. Adopting the standard precription for the
kinematic viscosity $\nu = \alpha c_s H$ \citep{shakura73}, where
$\alpha$ is the dimensionless viscosity parameter, 
the viscous timescale for a perturbation lengthscale 
$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
Setting $t_\mathrm{visc}$ to be longer than the
minimum VSI growth timescale $1/h\Omega$ implies 
$\khat^2\lesssim h/\alpha$. Inserting $h \sim 0.05$,
we find only modes with $\khat\lesssim 20$ are relevant for
$\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
10^{-5}$. \citepalias[These low viscosity levels are required for the
VSI, see][]{nelson13}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/bcrit_mink} 
  \caption{The minimum perturbation wavenumber $\khat$ in
    the MMSN such that the associated dimensionless thermal
    relaxation time $\beta$ at $z=0$ (solid) and $z=3H$ (dashed) is
    less than the critical timescale $\beta_\mathrm{crit}$.   
    \label{mmsn_bcrit_bcool_mink}}   
\end{figure}

\subsection{VSI in the MMSN}
Here we solve the linear stability problem in the MMSN with the scale and
height dependent thermal relaxation timescale given by
Eq. \ref{beta_mmsn_simp}. 
 Fig. \ref{beta_compare} show examples of
the thermal timescales at $r=5$AU and $r=50$AU for a
perturbation wavenumber $\khat=30$. We also plot the 
corresponding critical thermal timescales evaluated for the MMSN
(Eq. \ref{bcrit_mmsn}).  \\
\emph{For this figure to belong here (and not in the previous subsection, with the other comparisons of $\beta$ and $\beta_{crit}$) the discussion should be be more focused on the linear analysis.  Again it's better to start by discussing what the figure tells us, than repeating the figure caption. E.g.\ you could use Fig 21 to explain when a height varying $\beta$ is required.  I would also explain (especially if it stays in this subsection) what the growth time is for the r = 5 AU case.  Looks to be $\sim 1000$ orbits, which is remakrable given that $\beta < \beta_{\rm crit}$ outside $z \gtrsim  1.6 H$.?  That kind of detail could be illuminating, as it's nominally surprising given previous evidence that behavior near 5 -- 7 $H$ can affect growth.}

 \begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/beta_compare}
  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
    and $r=5$AU for $\khat=30$ (solid lines). The
    corresponding horizontal dashed lines are the critical thermal
    relaxation timescales derived in linear theory. 
    \label{beta_compare}}
\end{figure}

The outer disk ($r=50$AU) is optically thin,
$\beta<\beta_\mathrm{crit}$  and  does not depend on height.
Accordingly we expect, and find, the VSI operates at $r=50$AU. The
inner disk ($r=5$AU) is optically thick near the midplane but becomes
optically thin away from it, such that $\beta < \beta_\mathrm{crit}$
for $|z|\gtrsim1.25H$. In fact, we also find the VSI in this case, but
with much reduced growth rates (when scaled by the local rotation), as
described below.

In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
$t_\mathrm{grow} \equiv 1/\sigma$ of the fundamental mode with a range of
wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv
2\pi/\OmK$. In fact, we either only find the fundamental mode, or 
that it is the most unstable, except for the appearance of some 
surface modes for large wavenumbers $\khat\gtrsim 50$, which are 
boundary-dependent.  \\
\emph{above sentence too long and complex, could e.g.\ split off that surface modes are neglected and why.}
%However, surface modes are dependent on
%boundary conditions in a vertically isothermal disk \citep{barker15},  
%and 
Disturbances with large $\khat$ are also more likely subject to viscous 
decay, as discussed above.

% It is also unclear if surface modes are
% dynamically important given they are concentrated in the disk
% atmosphere where the density is small. 
% The dominance of the fundamental mode is also
% observed in hydrodynamic simulations of \cite{stoll14} which treat
% radiative cooling properly.   unclear if surface modes are
% dynamically important because the there is little mass in the disk at

We therefore consider Fig. \ref{mmsn_overall} as a representation of the
VSI in outer parts of the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
50$AU with growth timescales $\sim 30$---40 orbits and
$t_\mathrm{grow}$ rises rapidly towards the inner disk. It is
necessary to consider smaller scales with decreasing $r$ as the disk
generally becomes optically thick.

The fact that we obtain the VSI in the inner disk despite $\beta >
\beta_\mathrm{crit}$ near the midplane suggests that the instability
can operate provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
fraction of the vertical domain. However, the normalized growth times
become much longer. For example, for $\khat=30$ we find
$t_\mathrm{grow}\sim 40$ orbits at $r=50$AU, but $t_\mathrm{grow}\sim
400$ orbits at $r=5$AU. Thus, introducing an optically thick midplane
significantly reduces the efficiency of the VSI despite $\beta <
\beta_\mathrm{crit}$ away from the midplane.

%surface modes may develop for large kx, they are not affected

We thus find Fig. \ref{mmsn_overall} to be roughly consistent with
comparing the midplane thermal relaxation timescales to the critical
thermal timescale. For example, 
Fig. \ref{mmsn_bcrit_bcool} show that $\beta(z=0,\khat=10) \gtrsim
\beta_\mathrm{crit}$ for $r\lesssim 10$AU in the MMSN (left middle
panel); and Fig. \ref{mmsn_overall} 
show that modes with $\khat=10$ are stabilized inside $\sim 8$AU (where
$t_\mathrm{grow}\to\infty$). Our numerical results suggest the VSI
is efficient at radii of few tens of AU in the MMSN, consistent with
estimates made by \citetalias{nelson13}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
  \caption{Characteristic growth times of the VSI in 
    the MMSN. These are calculated using the thermal relaxation
    function given by Eq. \ref{beta_mmsn_simp}. 
    \label{mmsn_overall}}    
\end{figure}